{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-10-13 sftrends_manu.tex
figs tabs updated with new results, different seasonal comps

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Wed Jul 20 11:11:39 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

{\centering \includegraphics[width=0.65\textwidth]{figs/trends-1} 
\includegraphics[width=0.65\textwidth]{figs/trends-2}

@@

\caption[Percent changes in nitrogen analytes for (a) annual and (b) seasonal periods in the record]{Percent changes in nitrogen analytes for (a) annual and (b) seasonal periods in the record.  Changes are based on the difference between the ending and starting estimates for the flow-normalized estimates within each period.  Points are colored for direction (red increasing, green decreasing) and sized for relative magnitude. Station names are shown in the top left panel.}\label{fig:trends}
\end{figure}

@@

\caption[Nitrogen load measurements (kg d]{Nitrogen load measurements (kg d$^{-1}$) at the City of Tracy Wastewater Treatment Plant, San Joaquin County.  Wastewater discharge requirements were implemented in May, 2007 to include nitrification/denitrification and tertiary filtration causing a reduction in total nitrogen effluent discharged to the Delta. Reductions were primarily observed for ammonium.}\label{fig:tracy}
\end{figure}

@@

% trends in no23, nh at p8
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(2, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2), rowname = cats,     label = "tab:p8chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \ac{wwtp} upgrades upstream of station P8. Upgrades were completed in May 2007 at the City of Tracy \ac{wwtp} (San Joaquin County, \cref{fig:tracy}).  Summaries are means and percent changes based on annual means within the pre- and post-upgrade time periods (1976-2007, 2008-2013).  Increasing values are in bold-italics.\label{tab:p8chg}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NH$_{4}^{+}$}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries Annual}&&&&&\tabularnewline
~~1976-2007&1.27&\textit{\textbf{75.7}}&&0.33&\textit{\textbf{71.3}}\tabularnewline
~~2008-2013&1.47&-6&&0.11&-62.6\tabularnewline
\hline
{\bfseries Seasonal, pre}&&&&&\tabularnewline
~~JFM&1.46&\textit{\textbf{77.7}}&&0.49&\textit{\textbf{57.1}}\tabularnewline
~~AMJ&1&\textit{\textbf{75.3}}&&0.13&\textit{\textbf{63.4}}\tabularnewline
~~JAS&1.08&\textit{\textbf{97.9}}&&0.13&\textit{\textbf{151.9}}\tabularnewline
~~OND&1.53&\textit{\textbf{81.9}}&&0.54&\textit{\textbf{102.4}}\tabularnewline
\hline
{\bfseries Seasonal, post}&&&&&\tabularnewline
~~JFM&1.76&-3.4&&0.14&-47\tabularnewline
~~AMJ&0.94&-5.5&&0.07&-45.3\tabularnewline
~~JAS&1.17&-10&&0.07&-50.3\tabularnewline
~~OND&1.99&-4.6&&0.15&-58.5\tabularnewline
\hline
\end{tabular}\end{center}

@@

% trends in din, chl, sio2 at d7
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:d7chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg L$^{-1}$), chlorophyll ($\mu$g L$^{-1}$), and silicon dioxide (mg L$^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:d7chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&0.4&\textit{\textbf{27.8}}&&4.4&-81.6&&12.6&\textit{\textbf{30.2}}\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&0.4&\textit{\textbf{10.2}}&&9.8&-63.9&&10.8&\textit{\textbf{9}}\tabularnewline
~~1986-1994&0.4&\textit{\textbf{14.9}}&&3.4&-48.8&&12.2&\textit{\textbf{14.2}}\tabularnewline
~~1995-2003&0.5&\textit{\textbf{4}}&&2.1&-4.8&&13.5&\textit{\textbf{3.7}}\tabularnewline
~~2004-2013&0.5&-8.6&&2.5&\textit{\textbf{28.5}}&&13.7&-1.8\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&0.5&-4.9&&2.7&-56&&15&\textit{\textbf{4.4}}\tabularnewline
~~AMJ&0.4&\textit{\textbf{17.4}}&&6.1&-74.3&&12.9&\textit{\textbf{20.9}}\tabularnewline
~~JAS&0.3&\textit{\textbf{40.6}}&&6.2&-89&&10.9&\textit{\textbf{79.1}}\tabularnewline
~~OND&0.5&\textit{\textbf{13.1}}&&2.7&-83.8&&11.7&\textit{\textbf{33.7}}\tabularnewline
\hline
\end{tabular}\end{center}

@@

% trends in din, chl, sio2 at c10
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:c10chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station C10. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:c10chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&1.3&-16.5&&27.5&\textit{\textbf{9}}&&14.4&-21.7\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&1.3&\textit{\textbf{12.1}}&&26.1&-3.3&&15.7&-4.3\tabularnewline
~~1986-1994&1.4&\textit{\textbf{6.3}}&&27.5&\textit{\textbf{23.5}}&&14.8&-4\tabularnewline
~~1995-2003&1.4&-0.8&&29.3&-15.3&&14.1&-5\tabularnewline
~~2004-2013&1.2&-29.4&&27.1&\textit{\textbf{5.1}}&&13.1&-8.7\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&1.4&-20.5&&12&\textit{\textbf{14.8}}&&14.3&-16.9\tabularnewline
~~AMJ&1&-26.4&&43&\textit{\textbf{3.1}}&&13.1&-27.5\tabularnewline
~~JAS&1.3&-3.7&&46.6&-13.6&&14.8&-26.1\tabularnewline
~~OND&1.6&-6.8&&8.7&-18.5&&15.6&-16\tabularnewline
\hline
\end{tabular}\end{center}

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Wed Oct 12 11:34:29 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

{\centering \includegraphics[width=0.75\textwidth]{figs/trndmap-1} 
\includegraphics[width=0.75\textwidth]{figs/trndmap-2}

@@

\caption[Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record]{Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record.  Changes are based on the difference between the ending and starting estimates for the flow-normalized estimates within each period.  Points are colored for direction (red increasing, green decreasing) and sized for relative magnitude. Station names are shown in the top left panel. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.}\label{fig:trndmap}
\end{figure}

@@

\caption[Nitrogen load measurements (kg d$^{-1}$) at the City of Tracy Wastewater Treatment Plant, San Joaquin County]{Nitrogen load measurements (kg d$^{-1}$) at the City of Tracy Wastewater Treatment Plant, San Joaquin County.  Wastewater discharge requirements were implemented in May, 2007 to include nitrification/denitrification and tertiary filtration causing a reduction in total nitrogen effluent discharged to the Delta. Reductions were primarily observed for ammonium.}\label{fig:tracy}
\end{figure}

@@

% overall trends
%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = c(9,         9, 9), cgroup = c("Annual", "Seasonal"), n.cgroup = c(2,         4), rowname = stats, label = "tab:trnds")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and different time periods.  Summaries are averages (mg L$^{-1}$) and percent changes in parentheses (increasing in bold-italic).  Changes are based on the difference between the ending and starting estimates within each period.  See \cref{fig:trndmap} for a summary of spatial trends. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.\label{tab:trnds}} 
\begin{center}
\begin{tabular}{lllcllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Analyte/Station}&\multicolumn{2}{c}{\bfseries Annual}&\multicolumn{1}{c}{\bfseries }&\multicolumn{4}{c}{\bfseries Seasonal}\tabularnewline
\cline{2-3} \cline{5-8}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{1976-1995}&\multicolumn{1}{c}{1996-2014}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Spring}&\multicolumn{1}{c}{Summer}&\multicolumn{1}{c}{Fall}&\multicolumn{1}{c}{Winter}\tabularnewline
\hline
{\bfseries DIN}&&&&&&&\tabularnewline
~~C10&1.3 \footnotesize{(\textit{\textbf{35}})}&1.4 \footnotesize{(-28.6)}&&1.1 \footnotesize{(-25.8)}&1.2 \footnotesize{(-6.3)}&1.5 \footnotesize{(\textit{\textbf{5.5}})}&1.7 \footnotesize{(-20.7)}\tabularnewline
~~C3&0.4 \footnotesize{(\textit{\textbf{51.1}})}&0.5 \footnotesize{(\textit{\textbf{3}})}&&0.4 \footnotesize{(\textit{\textbf{62.2}})}&0.3 \footnotesize{(\textit{\textbf{71.9}})}&0.5 \footnotesize{(\textit{\textbf{54.4}})}&0.5 \footnotesize{(\textit{\textbf{18.3}})}\tabularnewline
~~D19&0.5 \footnotesize{(-5.5)}&0.4 \footnotesize{(-30.7)}&&0.5 \footnotesize{(-33.9)}&0.3 \footnotesize{(-25.4)}&0.3 \footnotesize{(-21.7)}&0.7 \footnotesize{(-33.5)}\tabularnewline
~~D26&0.4 \footnotesize{(\textit{\textbf{0.3}})}&0.5 \footnotesize{(-19.6)}&&0.5 \footnotesize{(-19.2)}&0.3 \footnotesize{(-10.2)}&0.4 \footnotesize{(-3.5)}&0.6 \footnotesize{(-21.7)}\tabularnewline
~~D28&0.5 \footnotesize{(-3.1)}&0.4 \footnotesize{(-45.6)}&&0.5 \footnotesize{(-32.9)}&0.2 \footnotesize{(-56.5)}&0.3 \footnotesize{(-57.4)}&0.8 \footnotesize{(-40.2)}\tabularnewline
~~D4&0.4 \footnotesize{(-8.1)}&0.4 \footnotesize{(-7.5)}&&0.4 \footnotesize{(-13.6)}&0.3 \footnotesize{(\textit{\textbf{4.3}})}&0.4 \footnotesize{(-1.2)}&0.5 \footnotesize{(-15)}\tabularnewline
~~D6&0.4 \footnotesize{(\textit{\textbf{11.6}})}&0.5 \footnotesize{(-6.1)}&&0.5 \footnotesize{(-5)}&0.4 \footnotesize{(\textit{\textbf{64.1}})}&0.4 \footnotesize{(\textit{\textbf{21}})}&0.5 \footnotesize{(-14.6)}\tabularnewline
~~D7&0.4 \footnotesize{(-1.7)}&0.5 \footnotesize{(-11.8)}&&0.5 \footnotesize{(-23.8)}&0.4 \footnotesize{(\textit{\textbf{49.9}})}&0.4 \footnotesize{(\textit{\textbf{10.4}})}&0.6 \footnotesize{(-34.6)}\tabularnewline
~~P8&1.4 \footnotesize{(\textit{\textbf{81.1}})}&1.8 \footnotesize{(-15.8)}&&1.5 \footnotesize{(\textit{\textbf{35.5}})}&1 \footnotesize{(\textit{\textbf{42}})}&1.6 \footnotesize{(\textit{\textbf{55}})}&2.3 \footnotesize{(\textit{\textbf{38.7}})}\tabularnewline
\hline
{\bfseries NH$_{4}^{+}$}&&&&&&&\tabularnewline
~~C10&0.1 \footnotesize{(-37.7)}&0 \footnotesize{(-55.5)}&&0.1 \footnotesize{(-68.6)}&0 \footnotesize{(-85)}&0.1 \footnotesize{(-90.8)}&0.2 \footnotesize{(-78.8)}\tabularnewline
~~C3&0.2 \footnotesize{(\textit{\textbf{127}})}&0.3 \footnotesize{(\textit{\textbf{14.1}})}&&0.2 \footnotesize{(\textit{\textbf{165.1}})}&0.2 \footnotesize{(\textit{\textbf{133.3}})}&0.3 \footnotesize{(\textit{\textbf{107.7}})}&0.2 \footnotesize{(\textit{\textbf{81.2}})}\tabularnewline
~~D19&0.1 \footnotesize{(\textit{\textbf{35.6}})}&0.1 \footnotesize{(-32.6)}&&0.1 \footnotesize{(-32.9)}&0 \footnotesize{(-21)}&0 \footnotesize{(-6.4)}&0.1 \footnotesize{(-12)}\tabularnewline
~~D26&0.1 \footnotesize{(\textit{\textbf{58.4}})}&0.1 \footnotesize{(-18.7)}&&0.1 \footnotesize{(\textit{\textbf{3.3}})}&0.1 \footnotesize{(\textit{\textbf{38.1}})}&0.1 \footnotesize{(\textit{\textbf{21.1}})}&0.1 \footnotesize{(\textit{\textbf{0.8}})}\tabularnewline
~~D28&0.1 \footnotesize{(-10)}&0 \footnotesize{(-31.8)}&&0.1 \footnotesize{(-42.8)}&0 \footnotesize{(-36.9)}&0 \footnotesize{(-42.2)}&0.1 \footnotesize{(-35.7)}\tabularnewline
~~D4&0.1 \footnotesize{(\textit{\textbf{74.6}})}&0.1 \footnotesize{(\textit{\textbf{0.3}})}&&0.1 \footnotesize{(\textit{\textbf{54.2}})}&0 \footnotesize{(\textit{\textbf{30.8}})}&0.1 \footnotesize{(\textit{\textbf{23.8}})}&0.1 \footnotesize{(\textit{\textbf{56.9}})}\tabularnewline
~~D6&0.1 \footnotesize{(\textit{\textbf{100.2}})}&0.1 \footnotesize{(-3.5)}&&0.1 \footnotesize{(\textit{\textbf{42.2}})}&0.1 \footnotesize{(\textit{\textbf{74.1}})}&0.1 \footnotesize{(\textit{\textbf{49.2}})}&0.1 \footnotesize{(\textit{\textbf{39.8}})}\tabularnewline
~~D7&0.1 \footnotesize{(\textit{\textbf{88.2}})}&0.1 \footnotesize{(-13.3)}&&0.1 \footnotesize{(\textit{\textbf{26.8}})}&0 \footnotesize{(\textit{\textbf{25.1}})}&0.1 \footnotesize{(\textit{\textbf{4.3}})}&0.1 \footnotesize{(\textit{\textbf{16.8}})}\tabularnewline
~~P8&0.3 \footnotesize{(\textit{\textbf{267.4}})}&0.3 \footnotesize{(-85.1)}&&0.2 \footnotesize{(-51.8)}&0.1 \footnotesize{(-12)}&0.3 \footnotesize{(-44.5)}&0.6 \footnotesize{(-61.2)}\tabularnewline
\hline
{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&&&&&&&\tabularnewline
~~C10&1.2 \footnotesize{(\textit{\textbf{42.8}})}&1.3 \footnotesize{(-26.9)}&&1.1 \footnotesize{(-21)}&1.2 \footnotesize{(-2.3)}&1.4 \footnotesize{(\textit{\textbf{23.3}})}&1.5 \footnotesize{(-5.6)}\tabularnewline
~~C3&0.2 \footnotesize{(\textit{\textbf{0.7}})}&0.2 \footnotesize{(-6.6)}&&0.2 \footnotesize{(\textit{\textbf{0.9}})}&0.1 \footnotesize{(\textit{\textbf{18}})}&0.1 \footnotesize{(\textit{\textbf{4.3}})}&0.2 \footnotesize{(-17.4)}\tabularnewline
~~D19&0.4 \footnotesize{(-11.7)}&0.4 \footnotesize{(-31.1)}&&0.4 \footnotesize{(-32.3)}&0.2 \footnotesize{(-36.7)}&0.3 \footnotesize{(-31.6)}&0.6 \footnotesize{(-38.4)}\tabularnewline
~~D26&0.3 \footnotesize{(-9.8)}&0.4 \footnotesize{(-20.1)}&&0.4 \footnotesize{(-23.5)}&0.2 \footnotesize{(-18.7)}&0.3 \footnotesize{(-10.5)}&0.5 \footnotesize{(-26.4)}\tabularnewline
~~D28&0.4 \footnotesize{(-10.5)}&0.4 \footnotesize{(-47.2)}&&0.5 \footnotesize{(-34.6)}&0.2 \footnotesize{(-66.3)}&0.3 \footnotesize{(-64.4)}&0.7 \footnotesize{(-45.7)}\tabularnewline
~~D4&0.3 \footnotesize{(-14.6)}&0.3 \footnotesize{(-9.8)}&&0.3 \footnotesize{(-19.2)}&0.3 \footnotesize{(\textit{\textbf{1.1}})}&0.3 \footnotesize{(-8)}&0.4 \footnotesize{(-25.9)}\tabularnewline
~~D6&0.3 \footnotesize{(-5.6)}&0.4 \footnotesize{(-8)}&&0.4 \footnotesize{(-14.5)}&0.3 \footnotesize{(\textit{\textbf{38.8}})}&0.3 \footnotesize{(\textit{\textbf{4}})}&0.4 \footnotesize{(-23.2)}\tabularnewline
~~D7&0.4 \footnotesize{(-13.4)}&0.4 \footnotesize{(-13.6)}&&0.4 \footnotesize{(-27.3)}&0.3 \footnotesize{(\textit{\textbf{28.7}})}&0.4 \footnotesize{(-23.7)}&0.5 \footnotesize{(-46.8)}\tabularnewline
~~P8&1.1 \footnotesize{(\textit{\textbf{59.8}})}&1.5 \footnotesize{(\textit{\textbf{3.3}})}&&1.2 \footnotesize{(\textit{\textbf{53.9}})}&1 \footnotesize{(\textit{\textbf{38.8}})}&1.4 \footnotesize{(\textit{\textbf{58.3}})}&1.8 \footnotesize{(\textit{\textbf{62.3}})}\tabularnewline
\hline
\end{tabular}\end{center}

@@

% trends in no23, nh at p8
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(2, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2), rowname = cats,     label = "tab:p8chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \ac{wwtp} upgrades upstream of station P8. Upgrades were completed in May 2007 at the City of Tracy \ac{wwtp} (San Joaquin County, \cref{fig:tracy}).  Summaries are means and percent changes based on annual means within the pre- and post-upgrade time periods (1976-2007, 2008-2013).  Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.\label{tab:p8chg}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NH$_{4}^{+}$}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries Annual}&&&&&\tabularnewline
~~1976-2007&1.3&\textit{\textbf{74.3}}&&0.34&\textit{\textbf{93.1}}\tabularnewline
~~2008-2013&1.5&-7.3&&0.11&-62\tabularnewline
\hline
{\bfseries Seasonal, pre}&&&&&\tabularnewline
~~Spring&1.24&\textit{\textbf{79.5}}&&0.25&\textit{\textbf{57.1}}\tabularnewline
~~Summer&0.99&\textit{\textbf{95}}&&0.09&\textit{\textbf{143.5}}\tabularnewline
~~Fall&1.32&\textit{\textbf{77}}&&0.29&\textit{\textbf{137.7}}\tabularnewline
~~Winter&1.67&\textit{\textbf{69.2}}&&0.69&\textit{\textbf{72.5}}\tabularnewline
\hline
{\bfseries Seasonal, post}&&&&&\tabularnewline
~~Spring&1.27&\textit{\textbf{0.9}}&&0.09&-49.2\tabularnewline
~~Summer&0.92&-18.7&&0.06&-49.2\tabularnewline
~~Fall&1.58&-8.7&&0.1&-58.2\tabularnewline
~~Winter&2.17&-5.1&&0.18&-62.1\tabularnewline
\hline
\end{tabular}\end{center}

@@

% trends in din, chl, sio2 at d7
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:d7chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg L$^{-1}$), chlorophyll ($\mu$g L$^{-1}$), and silicon dioxide (mg L$^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.\label{tab:d7chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&0.4&-7.8&&4.1&-76.7&&12.7&\textit{\textbf{21.8}}\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&0.4&-22.5&&9.3&-57.1&&10.9&-0.1\tabularnewline
~~1986-1994&0.4&\textit{\textbf{29.4}}&&3&-61.7&&12.2&\textit{\textbf{18.6}}\tabularnewline
~~1995-2003&0.5&-0.1&&2&\textit{\textbf{21.1}}&&13.7&\textit{\textbf{5.2}}\tabularnewline
~~2004-2013&0.5&-11.9&&2.4&\textit{\textbf{34.2}}&&13.7&-1.7\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~Spring&0.5&-23.8&&4.2&-60.3&&14.5&\textit{\textbf{0.6}}\tabularnewline
~~Summer&0.4&\textit{\textbf{49.9}}&&6.4&-82.3&&11.2&\textit{\textbf{51.3}}\tabularnewline
~~Fall&0.4&\textit{\textbf{10.4}}&&3.7&-84.9&&11.1&\textit{\textbf{43.3}}\tabularnewline
~~Winter&0.6&-34.6&&2&-63.8&&14.2&\textit{\textbf{4.2}}\tabularnewline
\hline
\end{tabular}\end{center}

% trends in din, chl, sio2 at c10
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:c10chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station C10. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.\label{tab:c10chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&1.4&-10.1&&33.2&-20.6&&14.4&-19.1\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&1.3&\textit{\textbf{22.8}}&&27.5&-21.6&&15.8&-3.2\tabularnewline
~~1986-1994&1.4&\textit{\textbf{1.2}}&&35.3&\textit{\textbf{31.5}}&&14.8&-4.1\tabularnewline
~~1995-2003&1.6&\textit{\textbf{16.4}}&&41&-27.5&&14.2&-3.3\tabularnewline
~~2004-2013&1.3&-36.3&&30.2&-12.8&&13.2&-7.6\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~Spring&1.1&-25.8&&26.2&\textit{\textbf{3.1}}&&13.2&-20\tabularnewline
~~Summer&1.2&-6.3&&79.6&-23.2&&13.9&-25.4\tabularnewline
~~Fall&1.5&\textit{\textbf{5.5}}&&19&-39.8&&15.6&-18\tabularnewline
~~Winter&1.7&-20.7&&7.7&\textit{\textbf{7.3}}&&15.1&-14.1\tabularnewline
\hline
\end{tabular}\end{center}
